Forecasting the old-age dependency ratio to determine a sustainable pension age

### 2020-09-18 PensionAge.Rmd
Nearly there

---
title: "Forecasting the old-age dependency ratio to determine the best pension age"
author:
- familyname: Hyndman
  othernames: Rob J
  address: Department of Econometrics & Business Statistics\newline Monash University, Clayton VIC 3800, Australia
  email: Rob.Hyndman@monash.edu
  correspondingauthor: true
- familyname: Zeng
  othernames: Yijun
  address: Monash University
- familyname: Shang
  othernames: Han Lin
  address: Macquarie University
abstract: "We forecast the old-age dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old-age dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population. Our results suggest that the Australian pension age should be increased to 68 by 2028 and 69 by 2034 in order to maintain the old-age dependency ratio at 23%, just above the 2018 level. Our general approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries."
keywords: "coherent forecasts, demographic components, functional time series model, pension and insurance"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: pensionage.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
header-includes:
  - \usepackage{todonotes, xcolor}
---

@@

# Set up drake plan
source("_drake.R")

@@

# Load results
drake::loadd()

# Load additional packages for graphs and tables
library(ggplot2)
library(patchwork)
```

\todo[inline]{Adjust numbers in abstract as per Figure 9}

@@

Using these predicted population paths, we can calculate the average OADR and predict the levels of old-age dependency that correspond to different pension age schemes. By testing different pension age schemes, we aim to find the pension age with the desired OADR outcome. Therefore, other contributions of this research are (1) to examine the effect of adjusting pension ages by comparing OADR before and after adjustment; and (2) to suggest a suitable pension age scheme to ensure that the economic burden will remain relatively constant at the desired level in the future. It should be noted that the desired level has been set at 21%, the estimated level of OADR in 2018. Whether this OADR level is the most appropriate level for Australia remains to be examined and is not the focus of this paper. This paper concentrates on providing a framework for policymakers to find a target pension age for a given desired level of OADR.

@@

The remainder of this paper proceeds as follows. Section \@ref(sec:litreview) reviews key literature on population forecasting and the pension age in Australia. Section \@ref(sec:data) describes the data and modelling framework we use in this study. We show how our framework can be used to estimate a pension age scheme that leads to a target OADR in Section \@ref(sec:target). Finally, Section \@ref(sec:conclusions) gives concluding remarks as well as identifies the scope for future research.

@@

```{r changespension, echo=FALSE, fig.cap="Australian pension age policies.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
# Policies obtained from https://www.dss.gov.au/seniors/benefits-payments/age-pension
# Law: https://guides.dss.gov.au/guide-social-security-law/3/4/1/10
# History: https://nationalseniors.com.au/news/campaign/a-short-history-of-the-age-pension-section
pension_age %>%
  filter(Year >= 2014) %>%
  ggplot(aes(x = Year, y = Age, group = Policy, col = Policy)) +
    geom_line() +
    ylab("Minimum pension age") +
    scale_color_brewer(
      palette = "Dark2",
      breaks = c("Proposed", "Current", "Age65"),
      labels = c("Proposed further change", "Current policy", "Pension age at 65")
    )
```

@@

# Results

@@

@@

```{r componentsfertility, fig.cap="Fitted basis functions and coefficients for Australian age-specific fertility rates.", message=FALSE}
autoplot.fmforecast(fertf)
```

@@

```{r totalpopulationbyyears, fig.cap="Fifty-year forecasts of the total population for each sex, along with 80% prediction intervals.", fig.height=4}
history_pop %>%
  group_by(Sex, Year) %>%
  summarise(Population = sum(Population) / 1e3) %>%
  ggplot(aes(x = Year, y = Population, group = Sex)) +
    geom_line(size = 1) +
    ylab("Population (millions)") +
    ggtitle("Total population") +
    facet_grid(. ~ Sex) +
    geom_ribbon(aes(y = Mean, ymin = Lo80, ymax = Up80), fill = "#7d7def", data = future_tot_pop) +
    geom_line(data = future_tot_pop, col = "#0000cc", mapping = aes(y = Mean), size = 1)
```

```{r totalpopulationin2038, fig.cap="Forecast population pyramid for 2038, along with 80% pointwise prediction intervals. The actual pyramid for 2019 is shown using dashed lines.", warnings=FALSE}
pyramid(
  history_pop %>% filter(Year == max(Year)),
  future_pop %>% filter(Year == max(history_pop$Year) + 20)
)
```

@@

```{r OADRcurrent, fig.cap="Fifty-year forecasts of the OADR associated with a pension age equal to 65 (green), the current approved pension age scheme (orange), and the proposed scheme with the pension age rising to 70 in 2035 (purple). The actual OADRs for 1921--2019 are shown in black."}
OADR_future %>%
  ggplot(aes(x = Year, y = OADR)) +
    geom_hline(yintercept = target, col = "gray") +
    geom_ribbon(aes(ymin = Lo, ymax = Hi, fill = Policy, group = Policy),
      alpha = 0.5) +
    geom_line(aes(col = Policy, group = Policy), size = .75) +
    geom_line(col = "black", size = .75, data = OADR_history) +
    xlab("year") +
    ylab("ratio") +
    ggtitle("OADR forecasts with pension age=65") +
    scale_color_brewer(palette = "Dark2") +
    scale_fill_brewer(palette = "Dark2")
```

Once the pension age reaches 67, the OADR under the current policy grows at nearly the same rate as the OADR of a pension age equal of 65. This is consistent with the idea that there is a persistent positive trend in OADR associated with a constant pension age, due to low fertility rates and better health care. Even the proposal to raise the pension age incrementally to 70 by 2035 leads to a similar increasing trend beyond 2035. Figure \@ref(fig:OADRcurrent) demonstrates that the pension age scheme requires constant review to ensure a stable OADR over the long term.

# Sustainable pension age scheme

@@

The algorithm to find the upper and lower boundary of plausible pension age schemes is the same as above, but mean of $O_{T+H|T}$ is replaced by the 10% and 90% quantiles.

```{r comparisonofpensionages, fig.cap="The comparison between the target pension age scheme (in green) and real pension age scheme that has been set or proposed by the government, along with the 80% confidence interval of the target pension age shown in the green shade. The solid red line indicates the changes to the pension age from 65 to 67, announced in 2009."}
pension.fc <- structure(
  list(
    mean = ts(pension_age_optimal$Age, start = min(pension_age_optimal$Year)),
    lower = ts(as.matrix(pension_age_lower$Age), start = min(pension_age_optimal$Year)),
    upper = ts(as.matrix(pension_age_upper$Age), start = min(pension_age_optimal$Year)),
    level = 80
  ),
  class = "forecast"
)
ggplot() +
  autolayer(pension.fc, series = paste("Pension age to give OADR =", target)) +
  autolayer(
    ts(pension_age %>% filter(Policy == "Current", Year >= 1995) %>% pull(Age),
       start = 1995),
    series = "Approved change"
  ) +
  autolayer(
    ts(pension_age %>% filter(Policy == "Age65", Year >= 1995) %>% pull(Age),
      start = 1995),
    series = "Pension age at 65"
  ) +
  xlab("Year") +
  ylab("Minimum pension age")
```

```{r OADRwithoptimal, fig.cap="The mean forecast of OADR associated with target pension age scheme (listed in Table 1) shown in blue, along with the desired OADR level in red (i.e., mean forecast of OADR in 2018). The 80% prediction interval of the forecast of OADR is shown as blue shading."}
## plots of OADR with plausible pension age schemes
OADR_optimal %>%
  ggplot(aes(x = Year)) +
    geom_hline(yintercept = target, col = "red") +
    geom_ribbon(aes(ymin = Lo, ymax = Hi), fill = "#7d7def") +
    geom_line(aes(y = OADR), col = "#0000cc") +
    geom_line(data = OADR_history, aes(y = OADR)) +
    ggtitle("OADR forecast estimates for optimal pension ages") +
    ylab("Old age dependency ratio")
```

```{r OADRboundaries, fig.cap="The 80% prediction interval forecast of OADR (in blue) associated with the upper and lower boundary of plausible pension age schemes (listed in Table 2), along with desired OADR level (in red)."}
bind_rows(
    OADR_upper %>% mutate(Series = "Upper boundary"),
    OADR_lower %>% mutate(Series = "Lower boundary"),
  ) %>%
  ggplot(aes(x = Year)) +
    geom_hline(yintercept = target, col = "red") +
    geom_ribbon(aes(ymin = Lo, ymax = Hi), fill = "#7d7def") +
    geom_line(aes(y = OADR), col = "#0000cc") +
    geom_line(data = OADR_history, aes(y = OADR)) +
    facet_grid(. ~ Series) +
    ggtitle("OADR forecasts for boundary pension age schemes") +
    ylab("Old age dependency ratio")
```

Table 1 shows the target pension age scheme, which is properly defined in section ??. Figure \@ref(fig:OADRwithoptimal) provides a visible comparison of the desired OADR level of 21% and the mean forecast of the OADR corresponding to the target pension age. Note that the OADR associated with target pension age scheme has nearly overlapped the desired OADR level, which indicates that we have successfully found the pension age with stable OADR at the desired levels.

```{r pensionagetables, message=FALSE, eval=FALSE}
pension_age_optimal %>%
  mutate(
    Month = round((Age - trunc(Age)) * 12),
    Agestring = paste0(
      trunc(Age), "Y", if_else(Month > 0, paste0("+", Month, "M"), "")),
    Column = trunc((Year - 2019) / 10) + 1,
  ) %>%
  select(Year, Agestring, Column) %>%
  nest(data = c(Year = Year, Age = Agestring)) %>%
  pivot_wider(names_from = Column, values_from = data) %>%
  unnest(cols = c(`1`, `2`, `3`, `4`), names_repair = "unique") %>%
  knitr::kable(
    format = "latex",
    col.names = rep(c("Year", "Age"), 4),
    booktabs = TRUE,
    caption = "Fifty-year target pension age scheme with stable OADR around desired OADR. Y denotes years and M denotes months."
  )
```

@@

```{r ComparisonwithOADR22, fig.cap="The comparison between the target pension age scheme (in green) associated with desired level equal to current level plus $1%$ (i.e., 22%) and real pension age scheme that has been set or proposed by government, along with the 80% confidence interval of the target pension age shown in green shade."}

```

```{r OADR_realpensionage, fig.cap="The mean forecast of OADR associated with different pension age schemes under investigation, along with the desired OADR level in grey (i.e., mean forecast of OADR in 2014). The 80% prediction interval of the forecast of OADR is shown as shading."}
```

The upper and lower boundaries of the plausible pension age schemes (also called the 80 % confidence interval of the target pension age scheme in this paper) are reported in Table 2, and their corresponding OADR simulation results are shown in Figure \@ref(fig:OADRboundaries). It can be seen that the lower (upper) limit of the 80% prediction interval of OADR associated with the lower (upper) boundary of the 80% confidence interval pension age scheme is entirely stable around the desired level. Note the negative relationship between OADR and pension age. It suggests, on the one hand, that any pension age higher than the upper boundary in Table 2 is deemed to be undesirable because the age is too high to allow the age pension to provide enough welfare support or insurance to the aged. On the other hand, any pension age scheme lower than the lower boundary of the pension age reported in Table 2 will also be regarded as unreasonable, because it sacrifices the interests of the next generation to provide a generous age pension immediately.

Figure \@ref(fig:comparisonofpensionages) compares the target pension age scheme and its 80% confidence interval with the pension age scheme with proposed further change. It was found that the target pension age increases at a higher rate than the proposed new pension age scheme. Reported in Table 1, the target pension age is supposed to reach 67 by 2021 and 70 by 2032, while pension age has been proposed to change to 67 by 2023 and 70 by 2035. Moreover, the proposed pension age scheme is below the lower boundary of plausible pension age schemes, which means we are 80% sure that it will cause an OADR higher than the desired level. From this result, it is concluded that the government will need to adjust pension age slightly quicker than the proposed pace to ensure we have a stable OADR around the desired level at 21%.

Note that the proposed pension age scheme does not diverge too far from the plausible pension age schemes, suggesting proposed new pension age scheme may be appropriate if we can afford slightly higher OADR. Therefore, it is worth performing a sensitivity test on the desired level here.

Using the approach suggested in this paper, the target pension age associated with the desired OADR equal to current level plus 1% has been found. We have shown the comparison of it \todo{code are missing for two figures} and its 80% confidence level with a real pension age scheme in Figure \@ref(fig:ComparisonwithOADR22). When we allow 1% more in desired OADR, the real pension age becomes plausible. Figure \@ref(fig:OADR_realpensionage) shows simulated OADR associated with proposed pension age scheme^[Since we don't know the change rate set by government for 2023--2035, we assume it follows increase rate for 2017--2023, which increase half year pension age in every two years.]. It confirms that OADR is stable around a level higher than current level of OADR before 2035. This results suggest that pension age scheme with proposed change could be suitable if the government are targeting a slightly higher financial pressure than current level.

@@

The ageing of the population is expected to become a critical issue in Australia due to the increase in life expectancy and the retirement of baby boomers. The fiscal cost of the age pension is projected to sharply increase as a result. For the purpose of sustainability of the welfare system in Australia, the government has announced an increase in the pension age to 67 in 2023 and 70 in 2035 to reduce the financial burden on society.

@@

We undertake the approach based on assuming the desired level of OADR is at the current level of 21%. Using the historical data from 1950 to 2018, the forecast of future populations suggests a substantial increase in total population, and confirms the inherent ageing trend. It was found that pension age should not be fixed, but instead should grow at a rate that will ensure a stable OADR at a level which can be affordable to Australians of working age. Findings indicate that the pension age scheme with proposed change was not too high, as it is still lower than the target pension age scheme and outside its confidence interval. It was concluded that the pension age set by the government should be adjusted slight quicker to be in line with our target pension age, which ensures a stable OADR of around 21%.

The empirical analysis presented in this paper shows the adjustment of the pension age announced in 2009 was effective in reducing the financial burden to the desired level, which was set at 21%. However, a one-time adjustment is not sufficient to ensure a stable OADR in the long term, due to ongoing changes in the population age structure. Therefore, we recommend consistently reviewing the pension age according to the presented statistical approach. This approach has been used to calculate the target pension age from now until 2068 and its 80% confidence interval. The target pension age found in this research increases to 67 by 2023, 70 by 2032 and 75 by 2058. This generates a stable OADR of around 21%. The rate of growth in the pension age as proposed by the government in 2014 is slightly lower than the target pension age found in this study, which suggests that the government may need to slightly accelerate the growth of the pension age to maintain an OADR at the current level.

@@

\newpage

# Appendix: Evaluating forecast accuracy {#sec:accuracy}

Since all the results and implications are dependent on the forecast accuracy, it is important to evaluate the accuracy of our forecast. It is not valid to assess the model's fit in terms of historical data. We are concerned with how well a model performs on out-of-sample data, which is known but not used to fit the model. Therefore, we will use a cross-validation procedure [@fpp] to evaluate the accuracy of the models used in this paper.

The logic of a cross-validation test is to use a portion of available data for fitting, and use the remaining data to test. Suppose we need $k$ years' observations to produce a reliable forecast^[$k$ can't be too small], we are interested in the $H$ prediction horizon. Then we are allowed to repeat the test $T-H-k+1$ times, where $T$ is the number of total years of observation. The procedure is as follows:

Let $\omega=1$, we first select historical data for a time interval of $[\omega,\omega+1,\dots,\omega+k-1]$ as "in sample data" to fit the model, and then forecast $H$ years' OADR^[The OADR is generated based on a pension age equal to 65 for the purpose of test.]. Then, we can calculate the difference of forecasted OADR over the period $[\omega+k,\omega+k+1,\dots,\omega+k+H-1]$ from the real OADR for the same period, which is known as absolute error. We repeat the following steps for $\omega=1,2,\dots,T-k-H+1$. Then, we can summarize the information into a matrix as follows:
$$
  \text{\textbf{Absolute Error}} = [\text{AE}_{h,\omega}] =
    \begin{bmatrix}
      \text{AE}_{1,1} & \cdots\cdots & \text{AE}_{1,T-k-H+1} \\
      \vdots          & \ddots       & \vdots \\
      \text{AE}_{H,1} & \cdots\cdots & \text{AE}_{H,T-k-H+1}
    \end{bmatrix}
$$
where $h$ denotes for the forecast horizon and $\text{AE}_{h,\omega}=\left|\text{OADR}^{\text{mean}}_{h,\omega}-\text{OADR}^{\text{real}}_{h,\omega}\right|$. This process is also known as "rolling forecasting origin", because the "origin" ($k+\omega-1$) at which the forecast is based rolls forward in time. We take the average of each row to obtain an absolute mean error of $h$-step-ahead forecast of OADR where $h=1,2,\dots,H$.

In our sample, $T=2018-1950+1=69$ and we set $k=25$ and $H=25$ to perform the test. The mean absolute error of $h$-step-ahead forecast of OADR for $h=1,2,\dots,25$ is shown in Figure \@ref(fig:rollingtest). From this Figure, we observe that the mean absolute error is increasing as $h$ increases, which is reasonable as we have less relevant information to make a forecast for further apart OADR. Overall, the mean absolute error is small. For example, the mean absolute error for 25-step-ahead forecast is 1.6%, which is small if compared to the desired level of OADR, which is at 25%.

```{r rmseplot}
oadr_rolling_sim %>%
  group_by(id) %>%
  mutate(h = row_number()) %>%
  ungroup() %>%
  select(Year, h, OADR) %>%
  rename(Forecast = OADR) %>%
  left_join(OADR_history, by = "Year") %>%
  mutate(e = OADR - Forecast) %>%
  group_by(h) %>%
  summarise(RMSE = sqrt(mean(e^2))) %>%
  ggplot(aes(x = h, y = RMSE)) +
    geom_point()
```

It would normally be ideal for comparing the mean absolute error of the forecast from our model to other available models to measure the accuracy. For example, suppose we can take the population projections from ABS reports over the last few decades, their projection of population age structure reported in different years can be used to mimic the rolling origin test of OADR, because we now have the real OADR for those projected years. However, we are not able to perform it because full projection data is no longer available online and ABS projections have not used a consistent and replicate methodology.

>>>>

---
title: "Forecasting the old-age dependency ratio to determine a sustainable pension age"
author:
- familyname: Hyndman
  othernames: Rob J
  address: Department of Econometrics & Business Statistics\newline Monash University, Clayton VIC 3800, Australia
  email: Rob.Hyndman@monash.edu
  correspondingauthor: true
- familyname: Zeng
  othernames: Yijun
  address: Monash University
- familyname: Shang
  othernames: Han Lin
  address: Macquarie University
abstract: "We forecast the old-age dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old-age dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population. Our results suggest that the Australian pension age should be increased to 68 by 2030, 69 by 2036, and 70 by 2050, in order to maintain the old-age dependency ratio at 23%, just above the 2018 level. Our general approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries."
keywords: "coherent forecasts, demographic components, functional time series model, pension and insurance"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: pensionage.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
header-includes:
  - \usepackage{todonotes, xcolor}
---

@@

# Set up plan
source("_drake.R")

@@

loadd(oadr_target)
```

@@

Using these predicted population paths, we can calculate the average OADR and predict the levels of old-age dependency that correspond to different pension age schemes. By testing different pension age schemes, we aim to find the pension age with the desired OADR outcome. Therefore, other contributions of this research are (1) to examine the effect of adjusting pension ages by comparing OADR before and after adjustment; and (2) to suggest a suitable pension age scheme to ensure that the economic burden will remain relatively constant at the desired level in the future. It should be noted that the desired level has been set at `r 100*oadr_target`%, just above the level of OADR in 2018. Whether this OADR level is the most appropriate level for Australia remains to be examined and is not the focus of this paper. This paper concentrates on providing a framework for policymakers to find a target pension age for a given desired level of OADR.

@@

The remainder of this paper proceeds as follows. Section \@ref(sec:litreview) reviews key literature on population forecasting and the pension age in Australia. Section \@ref(sec:results) describes the data and modelling framework we use in this study. We show how our framework can be used to estimate a pension age scheme that leads to a target OADR in Section \@ref(sec:target). Finally, Section \@ref(sec:conclusions) gives concluding remarks as well as identifies the scope for future research.

@@

```{r changespension, echo=FALSE, fig.cap="Australian pension age policies.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
# Policies obtained from https://www.dss.gov.au/seniors/benefits-payments/age-pension
# Law: https://guides.dss.gov.au/guide-social-security-law/3/4/1/10
# History: https://nationalseniors.com.au/news/campaign/a-short-history-of-the-age-pension-section
loadd(pension_age_plot)
pension_age_plot
```

@@

# Application to Australia {#sec:results}

@@

```{r loadfertfit}
loadd(fert.fit)
```

@@

```{r componentsfertility, fig.cap="Fitted basis functions and coefficients for Australian age-specific fertility rates.", message=FALSE}
loadd(fertf_plot)
fertf_plot
```

@@

```{r totalpopulationbyyears, fig.cap="Fifty-year forecasts of the total population for each sex, along with 80% prediction intervals.", fig.height=4}
loadd(population_plot)
population_plot
```

```{r totalpopulationin2038, fig.cap="Forecast population pyramid for 2038, along with 80% pointwise prediction intervals. The actual pyramid for 2019 is shown using dashed lines.", warnings=FALSE}
loadd(pyramid_plot)
pyramid_plot
```

@@

```{r OADRcurrent, fig.cap="Fifty-year forecasts of the OADR associated with a pension age equal to 65 (green), the current approved pension age scheme (orange), and the proposed scheme with the pension age rising to 70 in 2035 (purple). The actual OADRs for 1921--2019 are shown in black."}
loadd(oadr_plots)
oadr_plots
```

Once the pension age reaches 67, the OADR under the current policy grows at nearly the same rate as the OADR of a pension age equal of 65. This is consistent with the idea that there is a persistent positive trend in OADR associated with a constant pension age, due to low fertility rates and better health care. Even the proposal to raise the pension age incrementally to 70 by 2035 leads to a similar increasing trend beyond 2035. This proposal also leads to a slow decline in the OADR over the next 15 years. Figure \@ref(fig:OADRcurrent) demonstrates that the pension age scheme requires constant review to ensure a stable OADR over the long term.

# Sustainable pension age schemes

@@

The algorithm to find the upper and lower boundary of plausible pension age schemes is the same as above, but the mean of $O_{T+H|T}$ is replaced by the 10% and 90% quantiles.

Figure \@ref(fig:comparisonofpensionages) shows (in blue) the pension scheme produced by this algorithm, along with 80% confidence intervals. In comparison, the previous (green), current (red) and proposed (orange) schemes are also shown.

```{r comparisonofpensionages, fig.cap="The target pension age scheme (in blue) along with the pension age schemes that have been set and proposed by the government. The blue shaded region shows 80% confidence intervals for the target pension age."}
loadd(proposed_pension_scheme_plot)
proposed_pension_scheme_plot
```

```{r OADRwithoptimal, fig.cap="The mean forecast and 80% prediction intervals for the OADR associated with the target pension age scheme shown in blue, along with the desired OADR level in gray."}
## plots of OADR with plausible pension age schemes
loadd(proposed_oadr_plot)
proposed_oadr_plot
```

```{r OADRboundaries, fig.cap="The 80% prediction intervals of OADR (in blue) associated with the upper and lower boundary of plausible pension age schemes, along with desired OADR level in gray.", fig.height=3}
loadd(oadr_upper_lower_plot)
oadr_upper_lower_plot
```

Figure \@ref(fig:OADRwithoptimal) shows the OADR forecasts as a result of using the target age pension scheme shown in Figure \@ref(fig:comparisonofpensionages). The target level of `r oadr_target*100`% and the mean forecast of the OADR corresponding to the target pension age almost overlap. Figure \@ref(fig:OADRboundaries) shows the OADR forecasts corresponding to the upper and lower limits of the 80% confidence intervals shown in Figure \@ref(fig:comparisonofpensionages), demonstrating that they correspond to having OADR forecast prediction interval boundaries that lie very close to the target OADR value.

```{r pensionagetables, message=FALSE, eval=FALSE}
loadd(pension_age_optimal)
pension_age_optimal %>%
  mutate(
    Month = round((Age - trunc(Age)) * 12),
    Agestring = paste0(
      trunc(Age), "Y", if_else(Month > 0, paste0("+", Month, "M"), "")),
    Column = trunc((Year - 2019) / 10) + 1,
  ) %>%
  select(Year, Agestring, Column) %>%
  nest(data = c(Year = Year, Age = Agestring)) %>%
  pivot_wider(names_from = Column, values_from = data) %>%
  unnest(cols = c(`1`, `2`, `3`, `4`), names_repair = "unique") %>%
  knitr::kable(
    format = "latex",
    col.names = rep(c("Year", "Age"), 4),
    booktabs = TRUE,
    caption = "Fifty-year target pension age scheme with stable OADR around desired OADR. Y denotes years and M denotes months."
  )
```

@@

Thus any pension age higher than the upper boundary shown in Figure \@ref(fig:comparisonofpensionages) will give an OADR value that is significantly lower than the target value, while any pension age scheme lower than the lower boundary of the pension age reported in Figure \@ref(fig:comparisonofpensionages) will give an OADR value that is significantly higher than the target value.

Figure \@ref(fig:comparisonofpensionages) also compares the target pension age scheme and its 80% confidence interval with the pension age scheme that has been proposed but not yet enacted. Clearly, it is outside the 80% confidence interval and so is too high for our targeted OADR value of `r oadr_target*100`%. Under our proposed scheme, the Australian pension age should be increased to 68 by 2030, 69 by 2036, and 70 by 2050, in order to maintain the old-age dependency ratio at `r oadr_target*100`%.

@@

The ageing of the population is an increasingly critical issue in Australia (and many other countries) due to the increase in life expectancy and the retirement of baby boomers. The fiscal cost of the age pension is projected to sharply increase as a result. For the purpose of sustainability of the welfare system in Australia, the government has announced an increase in the pension age to 67 in 2023, with a proposal to increase it to 70 by 2035 in order to reduce the financial burden on society. Our analysis suggests that this increase is too rapid unless the aim is to reduce the OADR to below 2018 values.

@@

We undertake the approach based on assuming the desired level of OADR is around the 2018 level of `r oadr_target*100`%. Using historical data from 1950 to 2019, the forecast of future populations suggests a substantial increase in total population, and confirms the inherent ageing trend. It was found that pension age should not be fixed, but instead should grow at a rate that will ensure a stable OADR at a level which can be affordable to Australians of working age. Findings indicate that the pension age scheme with proposed change was not too high, as it is still lower than the target pension age scheme and outside its confidence interval. It was concluded that the pension age set by the government should be adjusted slight quicker to be in line with our target pension age, which ensures a stable OADR of around `r oadr_target*100`%.

The empirical analysis presented in this paper shows the adjustment of the pension age announced in 2009 was effective in reducing the financial burden to the desired level, which was set at `r oadr_target*100`%. However, a one-time adjustment is not sufficient to ensure a stable OADR in the long term, due to ongoing changes in the population age structure. Therefore, we recommend consistently reviewing the pension age according to the presented statistical approach. This approach has been used to calculate the target pension age from now until 2068 and its 80% confidence interval. The target pension age found in this research increases to 67 by 2023, 70 by 2050. This generates a stable OADR of around 21%. The rate of growth in the pension age as proposed by the government in 2014 is slightly lower than the target pension age found in this study, which suggests that the government may need to slightly accelerate the growth of the pension age to maintain an OADR at the current level.

@@

\clearpage
